---
title: 'ESM 204 HW 4'
author: "Caroline Shepherd"
date: "5/23/2022"
output: 
  html_document:
    theme: cerulean
    highlight: monochrome
    code.folding: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
    number_sections: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}

# Attach packages

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(equatiomatic)
library(cowplot)

```

```{r}

# Read in the data for damages (cost in $) and warming (in degrees C)

damages <- read_csv(here("damages.csv")) %>% 
  clean_names()

warming <- read_csv(here("warming.csv")) %>% 
  clean_names %>% 
  select(year:warming_pulse) 

```

# Estimating damages based on changes in global temperature

The first, “damages.csv” provides new model estimates of the annual total damages from climate change at different levels of warming (in degrees C). The second dataset, “warming.csv” contains estimates of a baseline future climate trajectory (in degrees C) until 2100, and a second trajectory that adds a one-time pulse of CO2 today to the atmosphere. The pulse is 35 billion tons of carbon, which is roughly equal to annual global emissions. You can think of this as a “small” one-time pulse in carbon emissions.1 

# Question 1

Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data. 

```{r}

# Pulse = 35 billion tons of carbon 
pulse <- 35000000000

# Derive a quadratic equation by making a new column with warming squared
damages$warming2 <- damages$warming^2

# Create a linear regression model of original warming and the squared warming data
damages_lm <- lm(damages ~ 0 + warming + warming2, data = damages)
summary(damages_lm)

# Estimate demand function by creating quadratic equation using the linear regression model
damages_funct <- function(x){
  damages_lm$coefficients[1]*x+damages_lm$coefficients[2]*x^2
}

# Quadratic equation:
extract_eq(model = damages_lm, use_coefs = TRUE, coef_digits = 2)

# Plot estimated damage function, overlaid with a scatterplot of the underlying data
ggplot(data = damages, aes(x = warming, y = damages)) +
  geom_point(size = .75) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "top") +
  labs(x = "\nWarming (Degrees Celsius)", y = "Damages ($)\n", title = "Damages and Global Temperature Change") +
  stat_function(fun = damages_funct, color = "red")

```

# Question 2

Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: 
(1) damages over time without the pulse, 
(2) damages over time with the pulse, 
(3) the difference in damages over time that arises from the pulse, and 
(4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage). 

```{r}

# Create new columns using lm model to depict each scenario from question 2
warming_damages <- warming %>%
  mutate(damages_baseline = damages_funct(warming_baseline)) %>% # 1) changes over time w/o pulse
  mutate(damages_pulse = damages_funct(warming_pulse)) %>% # 2) damages over time w/ pulse
  mutate(damages_difference = damages_pulse - damages_baseline) %>% # 3) diff in damages arising from pulse
  mutate(damages_difference_co2 = damages_difference/pulse) # 4) diff in damages over time w/ pulse/ton C02

```

### Plot all scenarios

```{r}

# Scenario 1: changes over time w/o pulse
ggplot(data = warming_damages, aes(x = year, y = damages_baseline)) +
  geom_line(color = "darkseagreen2", size = 1.5) +
  theme_bw() +
  labs(x = "Year", y = "Baseline Scenario Damages (Cost in $)")

```

Figure 2.0: This figure depicts changes in damages and warming over time without the pulse input.

```{r}

# Scenario 2: damages over time w/ pulse
ggplot(data = warming_damages, aes(x = year, y = damages_pulse)) +
  geom_line(color = "coral", size = 1.5) +
  theme_bw() +
  labs(x = "Year", y = "Pulse Scenario Damages (Cost in $)")

```

Figure 2.1: This figure depicts damages from warming over time with the inclusion of the pulse input.

```{r}

# Scenario 3: difference in damages arising from pulse
ggplot(data = warming_damages, aes(x = year, y = damages_difference)) +
  geom_line(color = "deeppink3", size = 1.5) +
  theme_bw() +
  labs(x = "Year", y = "Difference in Damages Arising from Pulse (Cost in $)")

```

Figure 2.2: This figure depicts the difference in damages arising from the inclusion of the pulse input.

```{r}

# Scenario 4: the difference in damages over time from the pulse per ton of CO2
ggplot(data = warming_damages, aes(x = year, y = damages_difference_co2)) +
  geom_line(color = "cadetblue4", size = 1.5) +
  theme_bw() +
  labs(x = "Year", y = "Difference in Damages/Ton of CO2 (Cost in $)")

```

Figure 2.3: This figure depicts the difference in damages over time from the pulse/ton of C02, assuming that each ton of the pulse results in an equal amount of damage.


# Question 3

The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC. 

```{r}

baseline_year <- 2022

warming_damages_scc <- warming_damages %>%
  select(year, damages_difference_co2) %>%
  mutate(pv_0 = damages_difference_co2*((1+0.0)^(-(year-baseline_year)))) %>%
  mutate(pv_1 = damages_difference_co2*((1+0.01)^(-(year-baseline_year)))) %>%
  mutate(pv_2 = damages_difference_co2*((1+0.02)^(-(year-baseline_year)))) %>%
  mutate(pv_3 = damages_difference_co2*((1+0.03)^(-(year-baseline_year)))) %>%
  mutate(pv_4 = damages_difference_co2*((1+0.04)^(-(year-baseline_year)))) %>%
  mutate(pv_5 = damages_difference_co2*((1+0.05)^(-(year-baseline_year)))) %>%
  mutate(pv_6 = damages_difference_co2*((1+0.06)^(-(year-baseline_year))))

discount_rates <- c(0,1,2,3,4,5,6)

scc_calc <- c(sum(warming_damages_scc$pv_0),sum(warming_damages_scc$pv_1),
                       sum(warming_damages_scc$pv_2),sum(warming_damages_scc$pv_3),
                       sum(warming_damages_scc$pv_4),sum(warming_damages_scc$pv_5),
                       sum(warming_damages_scc$pv_6))
              
scc_df <- data.frame(discount_rates,
                     scc_calc)

ggplot(data = scc_df, aes(x = discount_rates, y = scc_calc)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(x = "Discount Rate (%)", y = "Social Cost of Carbon ($/ton of CO2)")

# scc_function <- function(discount_rate){
#   warming_damages_scc %>%
#     mutate(pv = damages_difference_co2/((1+discount_rate)^(-(year-baseline_year))))
#   present_value = sum(warming_damages_scc$pv)
#   return(present_value)
# }
```



```{r}
ggplot(data = scc_df, aes(x = discount_rates, y = scc_calc)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(x = "Discount Rate (%)", y = "Social Cost of Carbon ($/ton of CO2)") +
  geom_label(aes(x=3, label="Ramsey Rule: Social Cost of Carbon = $71.35", y=85), color="coral") +
  geom_point(aes(x = 2.1, y = 71.35), color = "coral", size = 3)
```

Figure 3. The figure above depicts the social cost of carbon ($) per ton of C02 emissions at discount rates from 0-6%. As discount rate increases, the social cost of carbon decreases. This is expected because a higher discount rate indicates that we value our money more today, or in other words, that we value our money in the future less. This results on us placing a lower value on the social cost of carbon because the future effects are less of a concern. 

# Question 4

The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation: r = ρ + ηg Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above. 


```{r}
#Use the ramsey rule to calculate the discount rate for SCC
discount_rate_ramsey <- (0.001+2*0.01)*100
warming_damages_ramsey <- warming_damages %>%
  select(year, damages_difference_co2) %>%
  mutate(pv_ramsey = damages_difference_co2*((1+(discount_rate_ramsey/100))^(-(year-baseline_year))))
scc_ramsey <- sum(warming_damages_ramsey$pv_ramsey)
```

When using the Ramsey Rule, the discount rate we should use to calculate the SCC would be r discount_rate_ramsey%, and the resulting SCC would be r round(scc_ramsey, 2) dollars. This point is marked on the plot above where the blue vertical line intercepts the curve.


# Question 5

Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts 
-If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).
-Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever. 
-Society is risk neutral 
-Use a discount rate of 2% 


```{r}
discount_rate_policy <- 0.02
#assume society is risk neutral
#Policy A
warming_damages_policy_a <- warming_damages %>%
  select(year, warming_baseline, damages_baseline) %>%
  mutate(damages_2 = damages_funct(1.5*warming_baseline)) %>%
  mutate(pv_baseline = damages_baseline*(1+discount_rate_policy)^(-(year-baseline_year))) %>%
  mutate(pv_2 = damages_2*(1+discount_rate_policy)^(-(year-baseline_year)))
p_a_1 <- 0.5
p_a_2 <- 0.5
ev_a_1 <- sum(warming_damages_policy_a$pv_baseline)
ev_a_2 <- sum(warming_damages_policy_a$pv_2)
ev_a <- p_a_1*ev_a_1+p_a_2*ev_a_2 #expected present value of damages up to 2100 policy A in dollars
#Policy B
warming_damages_policy_b <- warming_damages %>%
  select(year, warming_baseline) 
for (i in 30:79){
  warming_damages_policy_b$warming_baseline[i]=1.29
}
warming_damages_policy_b <- warming_damages_policy_b %>%
  mutate(damages_b = damages_funct(warming_baseline)) %>%
  mutate(pv_b = damages_b*(1+discount_rate_policy)^(-(year-baseline_year)))
p_b <- 1.0
ev_b <- sum(warming_damages_policy_b$pv_b)*p_b #expected present value of damages up to 2100 policy B in dollars
```

### Policy A: Business as Usual

The expected present value of damages up to 2100 of policy A is r round(ev_a,0)` dollars.

### Policy B: Take Immediate Action on Climate Change

The expected present value of damages up to 2100 of policy B is r round(ev_b,0)` dollars.

```{r}
policy_a_cost <- function(x){ev_a}
policy_b_cost <- function(x){ev_b+x}
costs <- c(0,1e15,2e15,3e15,4e15,5e15,6e15,7e15,8e15,9e15,10e15)
total_cost_a <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)
total_cost_b <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)
cost_df <- data.frame(costs,total_cost_a,total_cost_b)
cost_df <- cost_df %>%
  mutate(total_cost_a = policy_a_cost(costs)) %>%
  mutate(total_cost_b = policy_b_cost(costs))
costs_df_long <- cost_df %>%
  pivot_longer(cols = c(total_cost_a,total_cost_b), names_to = 'policy', values_to = 'total_cost') %>%
  mutate(policy = case_when(
    policy == "total_cost_a" ~ "Policy A",
    policy == "total_cost_b" ~ "Policy B"
  ))
ggplot(data = costs_df_long, aes(x = costs, y = total_cost)) +
  geom_line(aes(color = policy)) +
  theme_minimal() +
  labs(x = "Initial Cost of Policy B ($)", y = "Expected Value of Policy ($)",
       title = "Expected Value of Policy versus Cost of Implementing Policy B")
#find point where policy a cost = policy b cost using uniroot()
dif_costs_fun <- function(x){ev_a-(ev_b+x)}
equal_cost <- uniroot(dif_costs_fun, c(1e15,2.5e15))$root
```

Figure 4. The above plot shows the expected value of policies A and B as a function of the initial cost of implementing Policy B (Policy A has no initial costs). 

The two policies have the same total expected value when the initial cost of Policy B is r round(equal_cost,0)` dollars. Therefore, if Policy B costs less than this amount, Congress should undertake Policy B, and if Policy B costs more than this amount, Congress should undertake Policy A (assuming that cost is their only consideration).

# Question 6 

Now suppose that society is risk averse. How do you think this would change your advice to Congress? 

If society is risk averse, they are less likely to want to take the risk that Policy A results in the lower cost scenario, or in other words, they are more likely to prefer Policy B because it has guaranteed lower future damages than the high cost outcome from Policy A. Therefore, if we determined that society is risk averse, I would advise Congress to model society's utility function from given monetary costs, then plug in the estimated costs of policies A and B into the utility function, and then recalculate the expected costs (or damages) of each policy using these new expected utility values. After this, they should determine the cost of Policy B, add it to the expected costs (or damages) of Policy B and then determine which policy has the lowest expected costs (or damages) and choose that policy. 